%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{article}

%\usepackage{endnotes}

\usepackage{url}	
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{mathtools}
\usepackage{csquotes}	

\usepackage{fixltx2e}
 
\usepackage[style=authoryear,backend=bibtex]{biblatex}

\addbibresource{references}
 
%\setlength{\parindent}{4em}
%\setlength{\parskip}{1em}

\begin{document}

\section{Introduction}

This report draws on the knowledge and experience gained from development
of the \cite{astro} and \cite{ivoa} virtual observatory systems to answer
the following questions:

\begin{itemize}
    \item What would be required to implement a virtual observatory system
    for the TROPGLOBE research community, capable of supporting the science
    use-cases outlined on the \cite{trop} website.
    \item What components from the \cite{astro} and \cite{ivoa} projects
    would be appropriate to use in developing a virtual observatory system
    for the TROPGLOBE research community.
\end{itemize}

\section{The IVOA Virtual Observatory}

The \citetitle*{ivoa} (\cite{ivoa}) was formed in June 2002 with a mission to:
\begin{quote}
"facilitate the international coordination and collaboration necessary for
the development and deployment of the tools, systems and organizational
structures necessary to enable the international utilization of astronomical
archives as an integrated and interoperating virtual observatory."
\end{quote}

The \citetitle*{vo} (\cite{vo}) is the realization of the \cite{ivoa} vision
of an integrated and interoperating virtual observatory.
The work of the \cite{ivoa} focuses on the development of standards,
providing a forum for members to debate and agree the technical standards
that are needed to make the \cite{vo} possible.

The operational \cite{vo} itself is comprised of a global shared metadata
registry, the \cite{ivoa-reg}, and a number of individual data discovery
and data access services deployed at each of the participating institutes.
These components work together to present a uniform mechanism for discovering
and accessing data, irrespective of where it is physically located.

The \cite{vo} architecture and data discovery processes are very similar
to the \textit{'interconnected metadata collections'} approach described
in \citetitle{jones-2006} (\cite{jones-2006}):
\begin{quote}
".... a loosely structured collection of project-specific data sets
accompanied by structured metadata about each of the data sets."
\end{quote}

\begin{quote}
"Each of the data sets is stored in a manner that is opaque to the data
system in that the data themselves cannot be directly queried; rather,
the structured metadata describing the data is queried in order to locate
data sets of interest."
\end{quote}

\begin{quote}
"After data sets of interest are located, more detailed information .... can
be extracted from the metadata and used to load, query, and manipulate
individual data sets."
\end{quote}

\subsection{Example use case}

A useful way to illustrate how the data discovery process works in the VO
is to look at an example task such as selecting images covering a particular
region of the sky, in a particular wavelength e.g. infrared, visible light,
radio or x-ray.

\subsubsection{Service discovery}

The first step of the process is to identify the services that provide
access to the type of data we are looking for by querying the \cite{ivoa-reg}.

The \cite{ivoa-reg} is comprised of a number of small local registry
services, typically hosted at the participating institute level, working in
cooperation with a set of higher level global registry services hosted by a
few key institutes that aggregate the data from the smaller registries to
create a global searchable index of metadata describing all of the services
and datasets available in the \cite{vo}.

When a new service is deployed, part of the deployment process involves
registering the service with the local registry.
The local registry is then responsible for collecting and storing the
metadata that describes both the service itself and the datasets that it
provides access to.
Once the metadata for a service or dataset has been registered in a local
registry, it is automatically propagated up to the next level and replicated
between the global registries.

This makes it possible to access the metadata for all of the services
and datasets published in the \cite{vo} by querying any one of the global
registries.

The first step in fulfilling our example use case is to identify services that
contain the type of data we are looking for, in this case images, by querying
the \cite{ivoa-reg} for services that support the \citetitle*{ivoa-sia}
(\cite{ivoa-sia}) capability.

In addition to the technical details of services and their capabilities
the \cite{ivoa-reg} also contains details about the content of datasets,
including details of the wavelength(s) measured, e.g. infrared, visible,
radio or x-ray.

This allows us to refine our query to search for \cite{ivoa-sia} services
that contain images in a specific waveband, e.g. optical, infrared or x-ray.

The \cite{ivoa-reg} query returns a table of data, each row of which contains
information about a \cite{ivoa-sia} service that provides the type of data
we are interested in - images in a particular wavelength.

The \cite{vo} is itself an evolving system, building on the existing work
to add additional levels of integration as new features are added to the
IVOA specifications.

A recent addition to the list of \cite{ivoa} standards is the
\citetitle*{ivoa-moc} (\cite{ivoa-moc}) which allows \cite{ivoa-reg}
services to perform coarse grained region matches.

This will enable us to further refine our \cite{ivoa-reg} query to filter
for \cite{ivoa-sia} services that contained data in a particular region of
the sky.

\subsubsection{Data discovery}

The next stage of the process is to query each of the \cite{ivoa-sia}
services in the list to discover details about the individual images
available from that service.

\noindent
A \cite{ivoa-sia} service can handle queries that specifiy a particular
wavelength and a particular region of the sky:
\begin{itemize}
  \item \texttt{POS}  The positional region (ra, dec).
  \item \texttt{BAND} The energy interval (wavelength).
\end{itemize}

Each \cite{ivoa-sia} service returns a table of data, each row of which
contains metadata about an individual image. The details of the fields
in the image metadata are defined in the \citetitle*{ivoa-obscore}
(\cite{ivoa-obscore}) data model.

This demonstrates a core part of the \cite{ivoa} architecture, interoperable
services based on standard interfaces and data formats.

All of the \cite{ivoa-sia} services will return a standard response, which
makes it much easier to combine them to produce a global list of all the
images available within the whole VO that match our search criteria.

\noindent
The two key components of this are:
\begin{itemize}
    \item A standard interface for the global \cite{ivoa-reg} that uses a
    standard set of attributes to describe datasets and services.
    \item A standard interface for local \cite{ivoa-sia} data access services
    that uses a standard set of attributes to describe the available data
    products.
\end{itemize}

The separation between the initial service discovery query at the global
level followed by individual data discovery queries at the local level is
very similar to the stages described in \cite{jones-2006}:
\begin{enumerate}
  \item Querying the metadata to establish the location of suitable data.
  \item Querying the individual services to establish what the data is and
  how to access it.
\end{enumerate}

\section{Tropical forest science}

\subsection{Carbon density comparison}

We can compare the \cite{vo} data discovery process for astronomy data with
an example use case based on a recent study \citetitle*{mitchard-2014}
(\cite{mitchard-2014}), comparing remote sensing data from satellites with
ground plot data collected in the field.

The study compares two sets of remote sensing data,
from \citetitle*{nasa-jpl-carbon} (\cite{nasa-jpl-carbon})
\citetitle{saatchi-2011} (\cite{saatchi-2011}), and the \citetitle*{whrc}
(\cite{whrc}) \citetitle{baccini-2012} (\cite{baccini-2012}), with four
sets of ground plot data from the following sources:

\begin{itemize}
    \item\citetitle*{rainfor} (\cite{rainfor}) (\cite{peacock-2007})
    (\cite{malhi-2009}).
    \item\citetitle*{atdn} (\cite{atdn}).
    \item\citetitle*{team} (\cite{team}).
    \item\citetitle*{ppbio} (\cite{ppbio}) (\cite{pezzini-2012}).
\end{itemize}

\subsubsection{Remote sensing source data}

The paper does not give details of the data discovery and data access
methods used to access the primary remote sensing source data.
However, there are a number of data discovery tools available that enable
researchers to search for remote sensing data products such as satellite
images and radar scans.

Good examples of this type of tool are the \cite{usgs-explorer} and
\cite{usgs-glovis} tools provided by the \citetitle*{usgs} (\cite{usgs}).

\begin{quote}
"The USGS EarthExplorer ... provides users the ability to query, search,
and order satellite images, aerial photographs, and cartographic products
from several sources."
\end{quote}

\begin{quote}
"In addition to data from the Landsat missions and a variety of other data
providers, EE now provides access to MODIS land data products from the NASA
Terra and Aqua missions, and ASTER level-1B data products over the U.S. and
Territories from the NASA ASTER mission."
\end{quote}

\begin{quote}
"The USGS Global Visualization Viewer (GloVis) is an online search and order
tool for selected satellite data. The viewer allows access to all available
browse images from the Landsat 7 ETM+, Landsat 4/5 TM, Landsat 1-5 MSS,
EO-1 ALI, EO-1 Hyperion, MRLC, and Tri-Decadal data sets, as well as Aster
TIR, Aster VNIR and MODIS browse images from the DAAC inventory."
\end{quote}

The \cite{usgs} also provides large area composited mosaics generated from
\cite{landsat} data via the \cite{weld} project.

\begin{quote}
"The WELD data products are processed so users do not need to apply the
equations, spectral calibration coefficients, and solar information, needed
to convert Landsat digital numbers to reflectance and brightness temperature.
They are defined in the same coordinate system and align precisely, making
them simple to use for multi-temporal applications.
The products provide consistent data that can be used to derive higher-level
land cover as well as geo-physical and biophysical products for assessment
of surface dynamics and to study Earth system functioning."
\end{quote}

The \cite{usgs} also maintains a \citetitle*{usgs-lta} (\cite{usgs-lta})
of historical remote sensing data:

\begin{quote}
"The U.S. Geological Survey's (USGS) Long Term Archive (LTA) at the National
Center for Earth Resource Observations and Science (EROS) in Sioux Falls,
SD is one of the largest civilian remote sensing data archives."
\end{quote}

\begin{quote}
"Time series images are a valuable resource for scientists, disaster
managers, engineers, educators, and the general public. USGS EROS has
archived, managed, and preserved land remote sensing data for more than 35
years and is a leader in preserving land remote sensing imagery."
\end{quote}

However, all of these interfaces are based around human interaction. There
do not apper to be any machine readable (\cite{web-service}) data discovery
services for this type of remote sensing data.

\subsubsection{Carbon density maps}

A detailed description of the dataset produced by \citetitle*{nasa-jpl-carbon}
(\cite{nasa-jpl-carbon}) is available in the associated paper
(\cite{saatchi-2011}).

The paper, along with the additional supporting information available on
the \citetitle*{pnas} (\cite{pnas}) website, describes the main upstream
data sources and the methods applied.

However, details of the data sources, the instruments, target areas and date
ranges the data covers are not available in a \cite{machine-readable} format.

The carbon density dataset itself is available as \cite{format-geotiff}
files, with associated \cite{format-world} metadata, for download from the
\cite{nasa-jpl-carbon-ftp} site.


A detailed description of the dataset produced by the \citetitle*{whrc}
(\cite{whrc}) is available in the associated paper (\cite{baccini-2012}).

The paper, along with the additional supporting information available from
the \cite{journal-nature} website, describes the upstream data sources and
the methods applied.
However, details of the data sources, the instruments, target areas and
date ranges the data covers are not available in a machine readable format.

The carbon density dataset itself is available by request from the
\cite{whrc-data} website. Access to the data requires filling in a simple
web form declaring who you are and what you want to use the data for. On
submitting the webform, an automated email reply is generated containing
a URL to a \cite{format-zip} file on the \cite{whrc} website.

The \cite{format-zip} file referred to in the email contains the data as
\cite{format-geotiff} files, with associated \cite{format-world} metadata.

This is a fairly standard mechanism for providing access to research
data. However, it is not suitable as a \cite{machine-readable}
\cite{web-service} interface.

\subsubsection{Ground plot data}

The four sets of ground plot data from \cite{rainfor}, \cite{atdn},
\cite{team} and \cite{ppbio} were combined together in the \cite{forest-plots}
database.

Details of the design and capabilities of the \cite{forest-plots} system
is presented in \citetitle*{gonzalez-2011} (\cite{gonzalez-2011}).

\begin{quote}
"The ForestPlots.net web application was designed primarily as a repository
for long-term intact tropical forest inventory plots, where trees within
an area are individually identified, measured and tracked through time."
\end{quote}

Of the three sets of ground plot data, the data from \cite{rainfor} and
\cite{atdn} were already available in the \cite{forest-plots} database.

The plot data from the \cite{team} and \cite{ppbio} projects were downloaded
and imported into the \cite{forest-plots} database manually.

A permanent archive of the combined field plot data is stored
in the \cite{forest-plots} database as a publically available
dataset\footnote{\url{http://dx.doi.org/10.5521/FORESTPLOTS.NET/2014_1}}
and is available in the supporting information for the paper.

\subsubsection{\cite{term-agb} data}

The \cite{term-agb} data for the forest plots were calculated using a
\cite{comp-lang-sql} query provided by the \cite{forest-plots} system which
implements the tropical forest model described in \citetitle*{chave-2005}
(\cite{chave-2005}).
The results of the \cite{term-agb} calculation for each forest plot are
included in the combined field plot dataset stored in the \cite{forest-plots}
database.

The paper refers to a number of maps derived from the field plot data and
other sources which were generated as part of the analysis:

\begin{itemize}
    \item \cite{kriged} map of mean wood density ($\rho$).
    \item Ratio of diameter (D) to tree height (H) \cite{feldpausch-2012}.
    \item \cite{kriged} map of basal area.
    \item \cite{kriged} map of \cite{term-agb} using D and species-specific
    $\rho$, and a regional height model (K\textsubscript{DH$\rho$}).
    \item \cite{kriged} map of \cite{term-agb} using D and species-specific
    $\rho$, but a pan Amazonian height model (K\textsubscript{D$\rho$}).
    \item \cite{kriged} map of \cite{term-agb} using D, regional height
    models and $\rho$, but with $\rho$ fixed at 0.63 (K\textsubscript{DH}).
    \item \cite{kriged} map of \cite{term-agb} using D, pan-Amazonian height
    model, and $\rho$ fixed at 0.63 (K\textsubscript{D}).
\end{itemize}

\begin{itemize}
    \item \cite{term-agb} map from \cite{saatchi-2011} (RS1).
    \item \cite{term-agb} map from \cite{baccini-2012} (RS2).
    \item Difference between RS1 and K\textsubscript{DH$\rho$}.
    \item Difference between RS2 and K\textsubscript{DH$\rho$}.
    \item Difference between RS1 and RS2.
\end{itemize}

These derived datasets and maps are not available in the supporting
information for the paper.

The \cite{term-agb} data derived from two remote-sensing-derived maps,
\cite{saatchi-2011} and \cite{baccini-2012} are not available in the
supporting information for the paper.

\section{\cite{trop} requirements}

Based on the \cite{trop} use cases we have studied so far it is clear that
data discovery forms a significant part of the requirements for \cite{trop}.

\subsection{External data}

In many of the use cases a significant part of the source material for
the use case has come from outside the \cite{trop} community.

For example, both the \cite{saatchi-2011} and \cite{baccini-2012} datasets
used in the \cite{mitchard-2014} use case came from external data sources,
\citetitle*{nasa-jpl-carbon} and \citetitle*{whrc} respectively.

In the short-term, in order to make this type of external data available
as part of the \cite{trop} data discovery process, it will be necessary
for a member of the \cite{trop} community to register and curate the
\cite{trop} metadata describing the external data.

In the longer-term, the ideal solution would be to encourage external data
providers like \citetitle*{nasa-jpl-carbon} and \citetitle*{whrc} to join
the \cite{trop} community and participate in the development of the standards
and \cite{web-service} interfaces for data sharing and discovery.

It is worth noting that a number of \cite{nasa} projects are active members
of the \cite{ivoa}, participating in the working groups and conferences
and contributing towards developing the \cite{ivoa} standards.

\subsection{Internal data}

A number of the \cite{trop} use cases require access to data provided by
members of the \cite{trop} community. Promoting and facilitating this kind
of data sharing and re-use of results within the \cite{trop} community is
one of the key goals of the \cite{trop} project.

In order to support this activity, the \cite{trop} system needs to enable
individual members of the \cite{trop} community to publish metadata describing
their datasets in the \cite{trop} system.

Once this metadata is available within the \cite{trop} system, it enables
other members of the \cite{trop} community to discover and use the data as
source material for their own research.

\section{\cite{ivoa} software}

In both cases, the requirements for the data discovery process are that the
users are able to specify an area of interest and the type of data they are
interested in and then gradually narrow the search criteria in response to
the data discovery results until they find the data they are interested in.

Based on this outline we can begin to evaluate how well the \cite{ivoa}
and \cite{astro} software meets the \cite{trop} requirements and compare
this with equivalent \citetitle*{gis} (\cite{gis}) software available.

At first glance, the \cite{ivoa} and \cite{trop} data discovery processes
are very similar. Suggesting that the \cite{ivoa} and \cite{astro} software
should be a good fit for the \cite{trop} requirements.

However, there are a number of issues that may mean that the \cite{ivoa} and
\cite{astro} software are not the best solution for meeting the \cite{trop}
requirements.

\subsection{Data models}

One issue is that a significant part of the \cite{ivoa} metadata structure
relies on a number of domain specific astronomy terms and concepts, making
it difficult to apply the \cite{ivoa} metadata structure to data from a
different domain.

Although it would be possible to remove the domain specific terms and
concepts from the \cite{ivoa} data model and replace them with a set of
terms that were more suited to the \cite{trop} domain.
Doing this piece at a time, gradually evolving a new metadata data model
for the \cite{trop} project would be a non-trivial undertaking involving
a significant commitment of time and resources.

It is worth noting that the \cite{ivoa} \cite{ivoa-obscore} data model that
forms the basis of the \cite{ivoa} data discovery process is the result
of 10 years' work by the \cite{ivoa} working groups to define a common data
model for astronomy observations.
It would be likely to take a similar length of time for the \cite{trop}
community to develop an equivalent data model from scratch.

With this in mind, it would probably be more practical to base the \cite{trop}
metadata on existing data models and data description techniques that are
already in use within the \cite{trop} community or that has been developed
for domains similar to \cite{trop}.

Examples of these are the \citetitle*{format-world} and \citetitle*{eml}
(\cite{eml}) metadata formats.

\subsubsection{\citetitle*{format-world}}

The \citetitle*{format-world} format provides a simple way of annotating
an existing map or raster image with \cite{gis} location metadata.

The \cite{format-world} format consists of a plain text file format containing
details of the location, scale and rotation of a map or raster image.

Both of the \cite{saatchi-2011} and \cite{baccini-2012} remote sensing
datasets provide \cite{format-world} metadata using the \textit{example.tfw}
convention to associate the metadata with the \cite{format-geotiff} maps.

This is a simple example of an established convention within the \cite{gis}
community for linking \cite{gis} metadata to datasets or maps.

\subsubsection{\citetitle*{eml}}

\cite{eml} is a detailed set of
specifications for metadata describing ecological datasets, based on
work done by the Ecological Society of America and associated efforts
\citetitle*{michener-1997} (\cite{michener-1997}).

\cite{eml} is implemented as a series of \cite{format-xml} document types
that can be used to describe different aspects of an ecological dataset.

These are just two examples of metadata data models and data description
techniques that are already in use within the \cite{gis} and Ecology domains.

This highlights a significant opportunity for the \cite{trop} community to
identify the existing metadata datamodels and data description techniques
already in use by members of the \cite{trop} community and to work together
to define a common set of interoperable models and techniques that best
describe the data used by the \cite{trop} community.

\subsection{Data owners}

A second issue with the \cite{ivoa} and \cite{astro} metadata \cite{ivoa-reg}
and data discovery tools concerns the allocation of roles and responsibilities
for managing the metadata within the \cite{ivoa-reg}, and the way these
reflect the structure of the \cite{ivoa} and the members involved in
developing the \cite{ivoa-reg}.

Historically, the most active contributors to the development of the
\cite{ivoa} standards and the \cite{astro} software have been primary data
providers within the international and UK astronomy communities.
Many of these represent large scale data providers responsible for
publishing and curating primary science archives for telescope surveys or
satellite missions.

In the \cite{trop} domain these are equivalent to the upstream data
providers who publish the original satellite remote sensing data, such as the
\cite{landsat} data archive or the \citetitle*{usgs-lta} (\cite{usgs-lta})
of remote sensing data published by by the \cite{usgs}.

This has influenced the way that the \cite{ivoa} and \cite{astro} software
and services have been developed. In particular, the priority has been
to concentrate on providing tools and
services for publishing the large primary source datasets.

This emphasis on the larger data providers has meant that the curation of the
dataset metadata was seen as a system administrator role. As a result, many
of the current tools for managing and curating datasets are designed around a
single system administrator role managing the metadata for an entire service,
rather that individual researchers managing the metadata for their own data.

In contrast, in the \cite{trop} use cases the hope is that a significant
portion of the data in the system will be provided by and curated by
individual researchers or small research groups.
For example, the results and supplementary data for the (\cite{mitchard-2014})
paper would be published and curated by the members of the research team
themselves.

\section{Alternative software}

The design issues identified with the \cite{ivoa} software would not prevent
using it as the basis for developing the \cite{trop} system.

It should be possible to gradually replace the \cite{ivoa-obscore} data model
in the \cite{ivoa} software with a new data model designed for \cite{trop},
and it should be possible to develop a new user interface and permission
infrastructure to enable individual users to publish and curate their
own data.

However, bearing in mind the level of resources these changes would require,
it may be useful to compare the data discovery capabilities and features
provided by the \cite{ivoa} and \cite{astro} software with the equivalent
capabilities of some alternative systems which are designed to handle
geographical and ecological data.

\subsection{Global Index of Vegetation-Plot Databases}

The \citetitle*{givd} (\cite{givd}) system is a complex registry of metadata
describing databases of vegetation plot data from around the world.

The \cite{givd} system contains records for over 200 databases and 3 million
individual vegetation plots.

Three of the datasets used in our use cases are listed in the \cite{givd}
system:

\begin{itemize}
    \item \texttt{[GIVD:00-00-001]}\footnote{\url{http://www.givd.info/ID/00-00-001}} \cite{forest-plots}.
    \item \texttt{[GIVD:SA-BR-001]}\footnote{\url{http://www.givd.info/ID/SA-BR-001}} \cite{ppbio}.
    \item \texttt{[GIVD:00-00-002]}\footnote{\url{http://www.givd.info/ID/00-00-002}} \cite{team}.
\end{itemize}

In \citetitle*{dengler-2011} (\cite{dengler-2011}) the \cite{givd} project
team describe the system architecture and some plans for the future to
aggregate different types of data from external sources.

\begin{quote}
"Our longer-term vision is to develop GIVD in ways similar to Metacat
(Jones et al. 2006), so that, ultimately, users who query GIVD will not
only receive information on which databases contain data suitable for the
intended analyses, but they will also discover other data from distributed
databases, with GIVD acting as the central node."
\end{quote}

This is broadly similar to the \cite{vo} architecture of distributed datasets
and to the \textit{'interconnected metadata collections approach'} described
in \citetitle{jones-2006} (\cite{jones-2006}).

However, the current emphasis is focussed on providing a human interactive
search facility, with the \cite{givd} system acting as the central node.
The current plans do not include providing a machine readable
\cite{web-service} interface to enable \cite{givd} system itself to be used
as a component in a larger distributed system.

\subsection{\cite{ppbio} Information System}

In \citetitle*{pezzini-2012} (\cite{pezzini-2012}) the \cite{ppbio} team
describe the role of the data manager and the metadata collection processes
developed as part of the \cite{ppbio} Information System.

They also describe the transition from an initial flat file data storage
system, to a new system based on \cite{metacat}.

\begin{quote}
"To facilitate data searches, all the metadata were converted to XML, and
the PPBio has installed a METACAT server to integrate with the Knowledge
Network for Biocomplexity (KNB), a network which aims to assist ecological
and environmental research."
\end{quote}

This move towards open standards for both the metadata (\cite{eml}) and the
service interfaces (\cite{metacat}) enables the \cite{ppbio} Information
System to become part of a larger distributed system.

\subsection{\citetitle*{knb}}

The \citetitle*{knb} ({\cite{knb}}) is a network designed to:

\begin{quote}
    "facilitate ecological and environmental research"
\end{quote}

by enabling researchers to:

\begin{quote}
    "share, discover, access and interpret complex ecological data."
\end{quote}

The \cite{knb} system is based on a set of \cite{open-source} software and
standards developed and maintained as part of the \cite{knb} project:

\begin{itemize}
    \item The \cite{morpho} data management tools.
    \item The rDataONE R package for accessing \cite{data-one} repositories.
    \item The \cite{metacat} metadata database.
    \item The \citetitle*{eml} (\cite{eml}) metadata language.
\end{itemize}

\subsubsection{\citetitle*{metacat}}

\cite{metacat} is a data management tool that provides a repository for
managing data and metadata in a single system:

\begin{quote}
"Metacat is a repository for data and metadata (documentation about data)
that helps scientists find, understand and effectively use data sets they
manage or that have been created by others."
\end{quote}

\cite{metacat} uses the \cite{eml} metadata data model and vocabulary to
describe datasets in the network.

In some cases the \cite{metacat} system stores both the metadata and actual
data itself, e.g.
\citetitle*{dietze-2004}\footnote{\url{https://knb.ecoinformatics.org/#view/doi:10.5063/AA/mdietze.3.2}}.

In other cases the \cite{metacat} system only stores the metadata, referring
to data that is stored elsewhere, e.g.
\citetitle*{dietze-2004}\footnote{\url{https://knb.ecoinformatics.org/#view/doi:10.5063/AA/mdietze.3.2}}.

This would be a good match for the \cite{trop} requirements and use cases.

In some cases, the \cite{trop} system needs to store both the data and the
metadata, e.g the results and supplementary data for (\cite{mitchard-2014}).

In this example, a member of the team working on (\cite{mitchard-2014})
would store the results and supplementary data in the \cite{trop} system
together with the metadata describing their data.

In other cases, the \cite{trop} system would only store the metadata,
along with a refence to the data stored in an external system, e.g. the
(\cite{saatchi-2011}) and (\cite{baccini-2012}) source material datasets
used by (\cite{mitchard-2014}).

In this example, a member of the team working on (\cite{mitchard-2014})
would enter the metadata details of the (\cite{saatchi-2011}) and
(\cite{baccini-2012}) datasets into the \cite{trop} system, but
the actual datasets would remain with their original publishers,
\citetitle*{nasa-jpl-carbon} and \citetitle*{whrc} respectively.

All three datasets would appear to be within the same (virtual)
system. Enabling \cite{trop} users to discover the (\cite{saatchi-2011})
and (\cite{baccini-2012}) remote sensing datasets alongside the results
and all of the supplementary data for the (\cite{mitchard-2014}) paper.

\subsection{\citetitle*{data-one}}

\cite{metacat} and the \cite{knb} project is part of the \citetitle*{data-one} (\cite{data-one})
project.

The \cite{data-one} project is part of the \citetitle*{nsf} (\cite{nsf})
funded \citetitle*{data-net} (\cite{data-net}) programme to build an
infrastructire that provides open, persistent, robust, and secure access
to Earth observational data.
 
\begin{quote}
"The DataONE project is a collaboration among scientists, technologists,
librarians, and social scientists to build a robust, interoperable,
and sustainable system for preserving and accessing Earth observational
data at national and global scales. Supported by the U.S. National
Science Foundation, DataONE partners focus on technological, financial,
and organizational sustainability approaches to building a distributed
network of data repositories that are fully interoperable, even when those
repositories use divergent underlying software and support different data
and metadata content standards."
\end{quote}

The \cite{data-one} architecture is based on a set of top level
\textit{Coordinating Nodes} and individual \textit{Member Nodes} located
at each participating institute or organisation.

The top level \textit{Coordinating Nodes} provide a replicated cataloge of
the data in the \textit{Member Nodes}, enabling researchers to search for
and discover data across the whole network.

The individual \textit{Member Nodes} at each institute enable researchers
to publish data to the whole \cite{data-one} network.

This hierarchical structure is similar to the \cite{vo} architecture of a global
\cite{ivoa-reg} containing metadata describing datasets and services in the
\cite{vo} and the \textit{'interconnected metadata collections approach'}
described in \citetitle{jones-2006} (\cite{jones-2006}).

\subsection{\citetitle*{fgdc-geo}}

The \citetitle*{fgdc} (\cite{fgdc}) \citetitle*{fgdc-geo} is designed to :

\begin{quote}
"provide a suite of well-managed, highly available, and trusted geospatial 
data, services, and applications for use by Federal agencies-and their State,
local, Tribal, and regional partners."
\end{quote}

The \cite{fgdc-geo} system brings together metadata standards, software
and services that provide a set of features which are similar to those that
the \cite{trop} aims to provide.

\begin{itemize}
    \item Map Viewer.
    \item Trusted Datasets.
    \item Multiple Basemaps.
    \item Collaborative Groups.
    \item Editable Layers.
\end{itemize}

\subsection{\citetitle*{ckan}}

A key component of the \citetitle*{fgdc-geo} system is the \citetitle*{ckan}
(\cite{ckan}) service, which provides the main metadata and data repository
for the system.

Development of \cite{ckan} \textit{data management system} (DMS) is managed
by the \citetitle*{okfn} network.

\cite{ckan} is used to power official data portals by national and local
governments in the UK, Brazil, the Netherlands, Austria, the US.

Examples of science based CKAN sites:
\begin{itemize}
    \item \citetitle*{bris} Research Data Repository\footnote{\url{http://data.bris.ac.uk/data/}}.
    \item \citetitle*{drdsi}\footnote{\url{http://drdsi.jrc.ec.europa.eu/}}.
    \item \citetitle*{netl} Energy Data eXchange\footnote{\url{https://edx.netl.doe.gov/about}}.
    \item \citetitle*{ngds} \footnote{\url{http://geothermaldata.org/}}.
    \item \citetitle*{noaa} data catalog\footnote{\url{https://data.noaa.gov/dataset}}.
\end{itemize}

As with the \cite{metacat} system, \cite{ckan} is able to store the data
along with the metadata describing it, or just store the metadata about
a data resource held in an external system.

This matches the two use cases described above, where members of a research
team would store their results and supplementary data in the \cite{trop}
system together with the metadata describing the data, or they would enter
the metadata for an external dataset stored in remote repository.

As with the \cite{metacat} system, \cite{ckan} is designed to function
as a node in a federated network of services, using a metadata harvesting
mechanism to bring together metadata about resources in other nodes.

This model of using distributed federation of collaborating services
is similar to the \cite{ivoa} and \cite{astro} architecture and the
\textit{'interconnected metadata collections approach'} described in
\citetitle{jones-2006} (\cite{jones-2006}).

\section{\cite{ivoa} process}


** TODO ** Need more details here


The thing we want to keep from \cite{ivoa} and \cite{astro}.

\begin{itemize}
    \item Both scientists and engineers involved.
    \item Working groups looking at specific areas.
    \item Development based on open source model.
    \item Standards evolve through public discussion.
    \item Small meetings focussed on specific technical details.
    \item Large meetings focussed on interoperability and standards.
\end{itemize}

\section{Conclusion}

\subsection{Core software}

For reasons outline above, the \cite{astro} software may not be the best
solution for meeting the \cite{trop} requirements.
It may be more practical to base the \cite{trop} system on one of the
\cite{gis} based systems such as \cite{metacat} or \cite{ckan}.

Both \cite{metacat} and \cite{ckan} already provide support for the \cite{gis}
and \cite{eml} metadata models. In addition, both projects are actively
involved in large scale cross disciplinary data management projects.

\begin{itemize}
    \item \cite{metacat} is part of the \cite{data-one} project, which is
    specifically aimed at building the infrastructure for earth observational
    data.
    \item \cite{ckan} is used in a number of national \cite{data-gov}
    projects, in particular the \cite{fgdc} \cite{fgdc-geo} system which
    is specifically aimed at handling geospatial data.
\end{itemize}

A number of the \cite{trop} science cases are based on primary remote
sensing data from sources such as \cite{landsat} and on processed data
products from groups such as \cite{nasa-jpl-carbon} and \cite{whrc}.

Ideally, in the long term, the best way to support the \cite{trop}
science cases would be to encourage the external data providers to join
the \cite{trop} community and adopt the same metadata and service standards.
 
Basing the \cite{trop} system on software that is already being used by
established cross disciplinary projects that handle earth observational
and geospatial data and use the \cite{gis} and \cite{eml} metadata
models is likely to make it easier encourage data providers such as
\cite{nasa-jpl-carbon} and \cite{whrc} to participate.

\subsection{Project structure}

However, that said, we would strongly recomend adopting the organizational
structure and development processes from the \cite{ivoa} and \cite{astro}
projects.


The \cite{astro} project succeded in its goal of delivering a \cite{vo}
system for the UK astronomy community, and as one of the founding members
of the \cite{ivoa} played a significant part in establishing the \cite{ivoa}
structure and processes.

The fact that over 10 years after it was first established the the \cite{ivoa}
is still an active community of scientists and engineers busy working on
developing the next generation of servces and specifications is in no small
part due to the organizational structures and processes put inplace early
in the \cite{astro} and \cite{ivoa} projects.


\subsection{Next steps}


\appendix

\section{National \cite{open-data} projects using CKAN}

\begin{itemize}
    \item Argentina (\url{http://datospublicos.gob.ar/})
    \item Austria   (\url{https://www.data.gv.at/})
    \item Australia (\url{http://data.gov.au/})
    \item Canada (\url{http://open.canada.ca/en})
    \item European Union (\url{http://open-data.europa.eu/en/data/})
    \item Germany (\url{https://www.govdata.de/})
    \item Italy (\url{http://www.dati.gov.it/})
    \item Netherlands (\url{https://data.overheid.nl/})
    \item Norway (\url{http://data.norge.no/})
    \item Ireland (\url{http://data.gov.ie/})
    \item Romanina (\url{http://data.gov.ro/})
    \item Slovakia (\url{http://data.gov.sk/})
    \item Switzerland (\url{http://opendata.admin.ch/})
    \item Uganda (\url{http://www.data.ug/})
    \item UK (\url{http://data.gov.uk/})
    \item USA (\url{http://www.data.gov/})
\end{itemize}

%\bibliographystyle{plain}
%\bibliography{sample}

\printbibliography

\end{document}
